GEE models

### 2017-10-03 doc/esm.Rmd
UK spelling

Generalized estimating equations (GEE) is a technique similar to mixed effects 
modeling, except it calculates the marginal population estimates compared to the
subject-specific estimates in mixed effects models. GEE is well suited to and
commonly used in longitudinal cohort studies, especially given its capacity to
handle missed visits. GEE is able to handle longitudinal data due to the
estimating of a working correlation matrix, which was chosen based on
quasi-likelihood information criteria (QIC). The auto-regressive of order 1
(AR1) matrix was chosen for the GEE models as it had the best model fit accessed
using QIC, though other matrices (eg. exchangeable) had similar fit (data not
shown). One main advantage of using methods such as GEE is the inclusion of the
working correlation matrix that takes into account the inherent covariance
between values measured at different time points when computing model estimates
and standard errors. AR1 is the most appropriate correlation matrix for
longitudinal data since the correlation matrix assumes lower intercorrelation of
outcomes within an individual as time passes. For example, BMI measured at the
baseline visit would be more highly correlated to BMI measured at the 3rd year
visit than compared to the 6th year visit.

For the covariates, they were chosen based on previous literature, from 
directed acyclic graph (DAG) recommendations, and from QIC. The DAG
recommendations were obtained from using the DAGitty software,
http://dagitty.net/. DAG is a technique that allows for identifying potential
confounding causal pathways and to limit potential bias introduced from
adjusting for colliding variables. A collider variable is a variable that is
determined from two or more other variables, such that in causal graph modeling,
the arrows from the other variables appear to 'collide' into the variable. QIC 
is a technique that balances the goodness of fit of the model with its relative
complexity (i.e. number of covariates) against other models being compared. This
identifies a specific model from a number of models that is the 'best' fit and
least complex of all the models. QIC and other information criterion methods are
used in conjunction with other techniques such as literature and DAG based. The 
final GEE model selected as best fitting was negligibly different when using 
insulin sensitivity or beta-cell function as an outcome and as such the same 
covariates were chosen for models with each of the outcomes. The covariates
medication use, WC, ALT, and MET would be considered classical confounders while
TAG would be considered a mediating variable.

>>>>

Generalized estimating equations (GEE) is a technique similar to mixed effects 
modelling, except it calculates the marginal population estimates compared to the
subject-specific estimates in mixed effects models. GEE is well suited to and
commonly used in longitudinal cohort studies, especially given its capacity to
handle missed visits. GEE is able to handle longitudinal data due to the
estimating of a working correlation matrix, which was chosen based on
quasi-likelihood information criteria (QIC). The auto-regressive of order 1
(AR1) matrix was chosen for the GEE models as it had the best model fit accessed
using QIC, though other matrices (eg. exchangeable) had similar fit (data not
shown). One main advantage of using methods such as GEE is the inclusion of the
working correlation matrix that takes into account the inherent covariance
between values measured at different time points when computing model estimates
and standard errors. AR1 is the most appropriate correlation matrix for
longitudinal data since the correlation matrix assumes lower intercorrelation of
outcomes within an individual as time passes. For example, BMI measured at the
baseline visit would be more highly correlated to BMI measured at the 3rd year
visit than compared to the 6th year visit.

For the covariates, they were chosen based on previous literature, from 
directed acyclic graph (DAG) recommendations, and from QIC. The DAG
recommendations were obtained from using the DAGitty software,
http://dagitty.net/. DAG is a technique that allows for identifying potential
confounding causal pathways and to limit potential bias introduced from
adjusting for colliding variables. A collider variable is a variable that is
determined from two or more other variables, such that in causal graph modelling,
the arrows from the other variables appear to 'collide' into the variable. QIC 
is a technique that balances the goodness of fit of the model with its relative
complexity (i.e. number of covariates) against other models being compared. This
identifies a specific model from a number of models that is the 'best' fit and
least complex of all the models. QIC and other information criterion methods are
used in conjunction with other techniques such as literature and DAG based. The 
final GEE model selected as best fitting was negligibly different when using 
insulin sensitivity or beta-cell function as an outcome and as such the same 
covariates were chosen for models with each of the outcomes. The covariates
medication use, WC, ALT, and MET would be considered classical confounders while
TAG would be considered a mediating variable.
